---
title: "Review Exercise"
output: html_document
---

```{r}
library(tidyverse)
library(nycflights13)
```

##### 1. Make a table that describes each plane. It should have a column for tailnum, another column for average arrival delay, and another for the year the plane was manufactured.
``` {r}
data_1 <- planes %>% 
  left_join(flights,by="tailnum") %>%
  group_by(tailnum) %>%
  summarise(avg_arr_delay = mean(arr_delay)) %>%
  unpack() %>%
  left_join(planes,by="tailnum") %>%
  select(c("tailnum","avg_arr_delay","year"))

data_1
```

Top flights with delays!
``` {r}
data_1 <- data_1[order(-data_1$avg_arr_delay),]
data_1
```

##### 2. Make a table where each row is a day of the year. The first column is the date. The 2:4 columns give the number of (scheduled) departures from EWR, LGA, and JFK.
```{r}
data_2 <- flights %>% 
  mutate(date=as.Date(with(flights, paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  group_by(date,origin) %>%
  summarise(count=n()) %>%
  pivot_wider(names_from = origin,values_from = count) %>%
  unpack()
data_2
```

##### 3. Make a table where each row is a day of the year. Each destination airport is a column. The elements (day x destination) give the number of flights to that destination. What should NA’s be?
``` {r}
data_3 <- flights %>% 
  mutate(date=as.Date(with(flights, paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  group_by(date,dest) %>%
  summarise(count=n()) %>%
  pivot_wider(names_from = dest, values_from = count) %>%
  unpack()
data_3
```
The reason for NA's in some columns is that there would not have been any flights going to that destination on that day.