---
title: "In Class Exercise"
output: html_document
---

```{r}
library(tidyverse)
library(nycflights13)
```

##### 1. Make a table that describes each plane. It should have a column for tailnum, another column for average arrival delay, and another for the year the plane was manufactured.
``` {r}
data_1 <- planes %>% 
  left_join(flights,by="tailnum") %>%
  group_by(tailnum) %>%
  summarise(avg_arr_delay = mean(arr_delay)) %>%
  unpack() %>%
  left_join(planes,by="tailnum") %>%
  select(c("tailnum","avg_arr_delay","year"))

head(data_1,10)
```

Top flights with delays!
``` {r}
data_1 <- data_1[order(-data_1$avg_arr_delay),]
data_1
```

##### 2. Make a table where each row is a day of the year. The first column is the date. The 2:4 columns give the number of (scheduled) departures from EWR, LGA, and JFK.
```{r}
data_2 <- flights %>% 
  mutate(date=as.Date(with(flights, paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  group_by(date,origin) %>%
  summarise(count=n()) %>%
  pivot_wider(names_from = origin,values_from = count) %>%
  unpack()
head(data_2,10)
```

##### 3. Make a table where each row is a day of the year. Each destination airport is a column. The elements (day x destination) give the number of flights to that destination. What should NA’s be?
``` {r}
data_3 <- flights %>% 
  mutate(date=as.Date(with(flights, paste(year, month, day,sep="-")), "%Y-%m-%d")) %>%
  group_by(date,dest) %>%
  summarise(count=n()) %>%
  pivot_wider(names_from = dest, values_from = count)

data_3[is.na(data_3)] <- 0
head(data_3,10)
```
The reason for NA's in some columns is that there would not have been any flights going to that destination on that day.

##### 4. Of the planes built before 1980, what is the average flight distance?
```{r}
data_4 <- planes %>% 
  filter(year<1980) %>%
  left_join(flights,by="tailnum") %>%
  group_by(tailnum) %>%
  summarise(mean_distance = mean(distance))
head(data_4,10)
```

##### 5. Compute the average arrival delay for each destination airport. Plot this against the airports altitude.
```{r}
data_5 <- flights %>% 
  group_by(dest) %>%
  summarise(mean_arr_delay = mean(arr_delay)) %>%
  left_join(airports, by = c("dest"="faa")) %>%
  na.omit()
head(data_5)
```

##### Plotting the average arrival delay with the airports altitude we get,
``` {r}
ggplot(data_5,aes(x=alt,y=mean_arr_delay)) + geom_line() +ggtitle("Mean Arrival Delay vs Altitude of the airports")
```