---
title: "Mid Term"
author: "Siddhant Thakur"
output: html_document
---
  
#### 1.  Map the delay by destination.
  
  Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here's an easy way to draw a map of the United States.  You are welcome to use this code or some other code. 

```{r, message=FALSE}
library(tidyverse)
library(nycflights13)
library(maps)
```

You might want to use the size or colour of the points to display the average delay for each airport.

```{r, warning=FALSE}

#flights
#planes
#airlines
#airports
#weather

# put your answer here.

flights %>%
  group_by(dest) %>%
  summarise(mean_delay=mean(arr_delay,na.rm=TRUE)) %>%
  left_join(airports,by=c("dest"="faa"))%>%
  ggplot(aes(lon,lat)) +
  borders("state") +
  geom_point(aes(size = mean_delay, color=mean_delay,alpha=0.2)) +
  coord_quickmap()
```


#### 2.   Do planes trade ownership?

You might expect that there's an implicit relationship between plane and airline, because each plane is flown by a single airline. Explore this conjecture using data. (Let's assume that the tail number of a plane does not change.)

```{r}
# put your answer here. 

# this data contains all the trips made by the that plane and the airline controlling it
plane_carrier <-flights %>%
  group_by(tailnum,carrier) %>%
  summarise(trips_made=n()) %>%
  drop_na()

plane_carrier

# this data consists of the tailnums which had an ownership change
carrier_change <- plane_carrier %>%
  group_by(tailnum) %>%
  summarise(count=n()) %>%
  filter(count>1) %>% 
  unpack()

carrier_change
ev_and_9e <- plane_carrier %>% 
  filter(tailnum %in% c("N146PQ","N153PQ","N176PQ","N181PQ","N197PQ","N200PQ","N228PQ","N232PQ")) %>%
  pivot_wider(names_from = carrier,values_from = trips_made)
ev_and_9e

dl_and_fl <- plane_carrier %>% 
  filter(tailnum %in% c("N933AT","N935AT","N977AT","N978AT","N979AT","N981AT","N989AT","N990AT","N994AT")) %>%
  pivot_wider(names_from = carrier,values_from = trips_made)
dl_and_fl

filter(airlines, carrier %in% c("9E","EV","DL","FL"))
```

We see that there has been a trade of ownership of some planes between Endeavor Air Inc to ExpressJet Airlines Inc. and Delta AirLines to AirTran Airways Corporation or vice versa,


#### 3a.  Plane's average speed.

Notice that ```flights$air_time``` is in minutes.  Make a new column that is the air time in hours. 

```{r}
# put your answer here.
new_flights <- flights %>%
  mutate(air_time_hour=round(air_time/60,digits=2))

new_flights
```


#### 3b. Average speed

For each flight, compute the average speed of that flight (in miles per hour).  Then, for each plane, compute the average of those average speeds. Display it in a histogram. You can use a base R histogram ```hist``` or ggplot's ```geom_histogram```. 

```{r}
# put your answer here.  
plane_avg_speed <- new_flights %>%
  mutate(flight_speed = round(distance/air_time_hour,digits=2)) %>%
  group_by(tailnum) %>%
  summarise(mean_average_speed = mean(flight_speed,na.rm=TRUE))

plane_avg_speed

ggplot(plane_avg_speed,aes(mean_average_speed,color='red')) + 
  geom_histogram(binwidth = 1.5,na.rm=TRUE) + 
  labs(x = "Average speed of each plane",
       y = "Frequency",
       title = "Frequency plot for average speed of each plane",
       caption = "Data Source: nycflights13")
```




#### 4. Bonus

Make a table where each row is a destination, each column is a carrier, and each element is the number of times that the carrier has flown to that destination.  Ensure that you only count flights that arrived at the destination. 

```{r}
# put answer here
flights %>%
  drop_na() %>%
  group_by(dest,carrier) %>%
  summarise(trips_made=n()) %>%
  pivot_wider(names_from = carrier,values_from = trips_made)
```
